\textbf{Parallel dragonfly algorithm} \\[2ex] \large High Performance Computing for Data Science course 2024/2025

### 2025-08-30 text-paper/paper.tex
tables

After the end of all tasks it is possible to check how much gain in performance we have obtained.
To do this it is necessary to introduce some theoretical concepts that will give us the tools that
will give us a good metric system to evaluate our results.Speedup is defined as the ratio between the the execution times of the serial and parallel
implementations, so the formula will be equal to: $$Speedup = \frac{T_{serial}}{T_{parallel}}$$ 
from this definition it is obvious that the more the speedup is the more the parallel implementation
performs better. There is a particular case where the time execution of the denominator can be defined
as $T_{parallel}=T_{serial}/p$, where \textit{p} is the number of cores: this is called \textit{linear speedup},
and in this situation it perfectly match with the classic speedup coefficient
Efficiency is another important parameter, and it is defined by the ratio of the speedup and the 
number of cores, giving the following formula: $$E = \frac{Speedup}{p}$$
also in this case the efficiency decrease due to the increase of the number of cores. In case
of linear speedup we get that the efficiency is equal to the inverse of $p$
\subsection*{results}
\label{results}
Now it is possible to show results obtained from the cluster,
using the configuration with 4 cores and 1 thread will be considered
equal as the serial time. The reason for this approach is that a single core computation would require 1024 times the workload of the 1024 process version, but we calibrated the 1024 core version to take 3 minutes;
so it should take around $3*1024/4=768min \approx 13h $, computing single core would have taken 4 times that.
\begin{table}[H]
\centering
\tiny  % Use tiny font size for maximum space saving
\setlength{\tabcolsep}{3pt}  % Reduce column separation
\begin{tabular}{|c|*{10}{c}|}
  \hline
  T/P & 4 & 8 & 16 & 32 & 64 & 128 & 256 & 512 & 1024 & 2048 \\
  \hline
  1 & -- & 14260.97 & 8471.77 & 3521.04 & 1849.25 & 1157.88 & 483.40 & 255.68 & 193.13 & 118.08 \\
  2 & -- & 9659.36 & 5628.55 & 3259.06 & 2127.71 & 921.88 & 433.69 & 281.84 & 123.21 & -- \\
  4 & 12785.30 & 7745.36 & 4652.30 & 2398.08 & 1465.56 & 687.89 & 453.28 & 252.71 & -- & -- \\
  8 & 13781.07 & 11326.55 & 3800.54 & 2358.14 & 1308.72 & 722.74 & 597.09 & -- & -- & -- \\
  \hline
\end{tabular}
\end{table}

As we can see the speedup increase almost linearly with the amount of processes used, in the efficency table it is shown that the efficency stays close to 1, until 512 process, then it deeps down. Probably it happens because the distance of the nodes is increasing (maybe throught a slower switch).
Instead increasing thread parallelism is not as effective, and it seems to stagnate after 2 threads (but converging to 2)
The reason could be that mpi pings cpu cores and process executions, so it could be limited to two (becuase Simultaneous Multithreading is usually set to 2).

>>>>

After the end of all tasks it is possible to check how much gain in performance we have obtained.
To do this it is necessary to introduce some theoretical concepts that will give us the tools that
will give us a good metric system to evaluate our results.Speedup is defined as the ratio between the the execution times of the serial and parallel
implementations, so the formula will be equal to: $$Speedup = \frac{T_{serial}}{T_{parallel}}$$ 
from this definition it is obvious that the more the speedup is the more the parallel implementation
performs better. There is a particular case where the time execution of the denominator can be defined
as $T_{parallel}=T_{serial}/p$, where \textit{p} is the number of cores: this is called \textit{linear speedup},
and in this situation it perfectly match with the classic speedup coefficient
Efficiency is another important parameter, and it is defined by the ratio of the speedup and the 
number of cores, giving the following formula: $$E = \frac{Speedup}{p}$$
also in this case the efficiency decrease due to the increase of the number of cores. In case
of linear speedup we get that the efficiency is equal to the inverse of $p$
\subsection*{results}
\label{results}
Now it is possible to show results obtained from the cluster,
using the configuration with 4 cores and 1 thread will be considered
equal as the serial time. The reason for this approach is that a single core computation would require 1024 times the workload of the 1024 process version, but we calibrated the 1024 core version to take 3 minutes;
so it should take around $3*1024/4=768min \approx 13h $, computing single core would have taken 4 times that.
\begin{table}[H]
\centering
\tiny  
\setlength{\tabcolsep}{3pt}
\begin{tabular}{|c|*{10}{c}|}
  \hline
  T/P & 4 & 8 & 16 & 32 & 64 & 128 & 256 & 512 & 1024 & 2048 \\
  \hline
  1 & 27879.62 & 14531.01 & 7700.45 & 3628.50 & 1873.90 & 1110.29 & 551.18 & 268.84 & 168.71 & 100.21 \\
  2 & 33202.40 & 14243.06 & 5811.25 & 3274.74 & 1946.32 & 903.55 & 425.83 & 249.58 & 129.12 & -- \\
  4 & 13136.02 & 8843.29 & 4623.91 & 2419.19 & 1460.78 & 724.75 & 461.13 & 260.10 & -- & -- \\
  8 & 15110.34 & 11308.10 & 3850.66 & 2345.90 & 1281.55 & 737.66 & 541.53 & -- & -- & -- \\
  \hline
\end{tabular}
\caption{Time results}
\label{time_results}
\end{table}

\begin{table}[H]
\centering
\tiny
\setlength{\tabcolsep}{3pt}
\begin{tabular}{|c|*{10}{c}|}
  \hline
  T/P & 4 & 8 & 16 & 32 & 64 & 128 & 256 & 512 & 1024 & 2048 \\
  \hline
  1 & 1 & 1.92 & 3.62 & 7.68 & 14.88 & 25.11 & 50.58 & 103.70 & 165.25 & 278.21 \\
  2 & 0.84 & 1.96 & 4.80 & 8.51 & 14.32 & 30.86 & 65.47 & 111.71 & 215.93 & -- \\
  4 & 2.12 & 3.15 & 6.03 & 11.52 & 19.09 & 38.47 & 60.46 & 107.19 & -- & -- \\
  8 & 1.85 & 2.47 & 7.24 & 11.88 & 21.75 & 37.79 & 51.48 & -- & -- & -- \\
  \hline
\end{tabular}
\caption{Speedup results}
\label{speedup_results}
\end{table}
\begin{table}[H]
\centering
\tiny
\setlength{\tabcolsep}{3pt}
\begin{tabular}{|c|*{10}{c}|}
  \hline
  Efficiency & 4 & 8 & 16 & 32 & 64 & 128 & 256 & 512 & 1024 & 2048 \\
  \hline
  1 & 1 & 0.96 & 0.91 & 0.96 & 0.93 & 0.78 & 0.79 & 0.81 & 0.65 & 0.54 \\
  2 & 0.42 & 0.49 & 0.60 & 0.53 & 0.45 & 0.48 & 0.51 & 0.44 & 0.42 & -- \\
  4 & 0.53 & 0.39 & 0.38 & 0.36 & 0.30 & 0.30 & 0.24 & 0.21 & -- & -- \\
  8 & 0.23 & 0.15 & 0.23 & 0.19 & 0.17 & 0.15 & 0.10 & -- & -- & -- \\
  \hline
\end{tabular}
\caption{Efficiency results (normalized)}
\label{efficiency_results_normalized}
\end{table}

As we can see the speedup increase almost linearly with the amount of processes used, in the efficency table it is shown that the efficency stays close to 1, until 512 process, then it deeps down. Probably it happens because the distance of the nodes is increasing (maybe throught a slower switch).
Instead increasing thread parallelism is not as effective, and it seems to give a $13\% $ improvment passing from 1 thread to 2 threads, $ 49\%$ to 4 threads, and $ 53\%$ to 8 threads.
It seems that the thread level paralelization can be improved, however it seems that there is an hard limit, and we think it has to do with mpi core pinning.
However the mpi scaling can replace openmp even on the same machine, providing better paralelism.
